PANCHANGE_analysis

### 2020-12-03 PANCHANGE_analysis.Rmd

```{r gad.sum violin plot Sample, fig.height=14, fig.width=8}
dat.long.gad.sum %>%
  mutate(Time_point_renamed = fct_recode(Time_point, 
            "Prepandemic" = "prepan",
            "Retrospective" = "retro",
            "Baseline" = "base")
         ) %>%
         mutate(Time_point_reordered = fct_relevel(Time_point_renamed, 
            "Baseline",
            "Retrospective",
            "Prepandemic")
         ) %>%
  ggplot(mapping = aes(
    x = Time_point_reordered,
    y = gad.sum_score,
    fill = Time_point_reordered)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = 10, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = 15, linetype = "dashed", color = "gray") +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    trim = FALSE,
    alpha = 0.5
    ) +
  geom_point(
    aes(
      y = gad.sum_score,
      color = Time_point_reordered),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8) +
#  geom_boxplot(
#    width = .1,
#    guides = FALSE,
#    outlier.shape = NA,
#    alpha = 0.5) +
 stat_summary(fun.y = mean,
              fun.ymin = function(x) mean(x) - sd(x), 
              fun.ymax = function(x) mean(x) + sd(x), 
              geom = "pointrange"
              ) +
  labs(
    y = "GAD7 scores",
    fill = "Time point") +
  theme_personal +
  theme(
    axis.text.y = element_blank()
  ) +
  scale_fill_manual(
    values = COPINGpalette3,
    breaks = c("Prepandemic", "Retrospective", "Baseline")) +
  scale_alpha(guide = 'none') +
  scale_size(guide = 'none') +
  scale_color_manual(
    values = COPINGpalette3,
    breaks = c("Prepandemic", "Retrospective", "Baseline"),
    guide = "none") +
  facet_wrap(facets = vars(Sample),
             nrow = 4
             ) +
  coord_flip()

@@

```{r model1b function}
model1b.estimates <- function(outcome, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23"
                                )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

# Regression diagnostic statistics
gad_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "gad.sum_score_base",
                                                 data_set = "dat")
kable(gad_baseline_model1b_reg_stats)
gad_baseline_model1b_reg_stats
```

@@

### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed_reordered + Ethnicity_collapsed + Time registration and baseline
```{r gad model1b diff_score prepan}
# Regression coefficients
gad_prepan_model1b_estimates <- model1b.estimates(outcome = "gad.diff_score_base_prepan",
                  data_set = "dat.prepan")
kable(gad_prepan_model1b_estimates)
gad_prepan_model1b_estimates

# Regression diagnostic statistics
gad_prepan_model1b_reg_stats <- model1b.reg.stats(outcome = "gad.diff_score_base_prepan",
                  data_set = "dat.prepan")
kable(gad_prepan_model1b_reg_stats)
gad_prepan_model1b_reg_stats
```

>>>>

```{r gad.sum violin plot Sample, fig.height=14, fig.width=8, dpi=300}
dat.long.gad.sum %>%
  mutate(Time_point_renamed = fct_recode(Time_point, 
            "Prepandemic" = "prepan",
            "Retrospective" = "retro",
            "Baseline" = "base")
         ) %>%
         mutate(Time_point_reordered = fct_relevel(Time_point_renamed, 
            "Baseline",
            "Retrospective",
            "Prepandemic")
         ) %>%
  ggplot(mapping = aes(
    x = Time_point_reordered,
    y = gad.sum_score,
    fill = Time_point_reordered)) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = 10, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = 15, linetype = "dashed", color = "gray") +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    trim = FALSE,
    alpha = 0.5
    ) +
  geom_point(
    aes(
      y = gad.sum_score,
      color = Time_point_reordered),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8) +
#  geom_boxplot(
#    width = .1,
#    guides = FALSE,
#    outlier.shape = NA,
#    alpha = 0.5) +
 stat_summary(fun.y = mean,
              fun.ymin = function(x) mean(x) - sd(x), 
              fun.ymax = function(x) mean(x) + sd(x), 
              geom = "pointrange"
              ) +
  labs(
    y = "GAD7 scores",
    fill = "Time point") +
  theme_personal +
  theme(
    axis.text.y = element_blank()
  ) +
  scale_fill_manual(
    values = COPINGpalette3,
    breaks = c("Prepandemic", "Retrospective", "Baseline")) +
  scale_alpha(guide = 'none') +
  scale_size(guide = 'none') +
  scale_color_manual(
    values = COPINGpalette3,
    breaks = c("Prepandemic", "Retrospective", "Baseline"),
    guide = "none") +
  facet_wrap(facets = vars(Sample),
             nrow = 4
             ) +
  coord_flip()

@@

```{r model1b function}

# function to return standardised beta for plotting in a dataframe

model1b.lm <- function(outcome, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23"
                                )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

  # run the glm
  assign(paste0(outcome,".model1b.lm"),
         lm( formula = glm.formula,
             data = get(data_set)
         )
  )

  return(get(paste0(outcome,".model1b.lm")))
}

model1b.estimates <- function(outcome, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23"
                                )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

# Regression diagnostic statistics
gad_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "gad.sum_score_base",
                                                 data_set = "dat")
kable(gad_baseline_model1b_reg_stats)
gad_baseline_model1b_reg_stats

# get lm object to standardise beta

gad_prepan_model1b_lm <- model1b.lm(outcome = "gad.diff_score_base_prepan",
                                           data_set = "dat.prepan")

@@

### Model 1b: Overlapping disorders, Adjusted
Model 1b + Gender_collapsed + age_category_collapsed_reordered + Ethnicity_collapsed + Time registration and baseline
```{r gad model1b diff_score prepan}
# Regression coefficients
gad_prepan_model1b_estimates <- model1b.estimates(outcome = "gad.diff_score_base_prepan",
                  data_set = "dat.prepan")
kable(gad_prepan_model1b_estimates)
gad_prepan_model1b_estimates

# Regression diagnostic statistics
gad_prepan_model1b_reg_stats <- model1b.reg.stats(outcome = "gad.diff_score_base_prepan",
                  data_set = "dat.prepan")

kable(gad_prepan_model1b_reg_stats)
